### 2016-08-01 capability-machine.tex
Added well-definedness lemmas for the logical predicates.

@@

The ticket dispenser test program uses $\malloc$. To be able to reason about programs that use $\malloc$, we assume that it follows the following specification. If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory. 
\begin{specification}[Malloc v.2]
  \begin{align*}
    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments \ldotp \\
    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a,m_b,m_e \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
    &\qquad \exists i, \var{FS} \ldotp \\
    &\qquad \quad W (i) = \iota_{\malloc}^{\var{FS}} \land \\
    &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\
    &\qquad \quad \memheap = \hsfoot \uplus \hsframe \land\\
    &\qquad \quad \heapSat[\hsfoot]{n}{W} \land \\
    &\qquad \quad \memreg(r_1) = \size \land \\
    &\qquad \quad \size \geq 0 \land \\
    &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
    &\qquad \quad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\
    &\qquad \quad \Rightarrow \\
    &\qquad \qquad\exists \Phi' \in \Mems \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
    &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
    &\qquad \qquad \qquad \memheap[\Phi']=\hsfoot' \uplus \hsframe \land\\
    &\qquad \qquad \qquad W' \future W \land \\
    &\qquad \qquad \qquad W(i) = \iota_{\malloc}^{\var{FS} \setminus [b',e']} \land \\
    &\qquad \qquad \qquad [b',e'] \subseteq \var{FS}  \land \\
    &\qquad \qquad \qquad \heapSat[\hsfoot']{n-j}{W'} \land \\
    &\qquad \qquad \qquad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
    &\qquad \qquad \qquad \size = e'-b' \\
    &\qquad \qquad \qquad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
    &\qquad \qquad \qquad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
\end{align*}
\end{specification}
\lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depent on the implementation (this is mostly a note to myself.)}
\dominique{I agree with the remark about the pc.  Perhaps leave the form of the
  pc unspecified as it is not important for malloc users?}
\dominique{Resulting frame should be equal to original.}
\dominique{Allocated pointer should have the right size.}
\dominique{Other registers unmodified could be an alternative. (Caller-save/Callee-save etc.)}

@@

\section{Logical Relation}
\label{sec:logical-relation}
\subsection{Worlds}
\[
  \Worlds = \{ W \in RegionNames \parfun \Regions \mid \dom(W) \text{ finite } \}
\]
$\States$ is a fixed set with the necessary states.
l
\begin{align*}
  \Regions = \{ r = (s, \phi, H) \mid &  s \in \States \land \phi \subseteq States^2 \land \\ 
                                      &  H : \States \rightarrow \StorePred \land \\ 
                                      &  \phi \text{ reflexive and transitive} \}
\end{align*}
\[
  \StorePred = \widehat{\Worlds} \monnefun \UPred{\HeapSegments}
\]

\[
\roll : \frac{1}{2} \cdot \Worlds \cong \widehat{\Worlds}
\]
Future Worlds:
\[
 W' \future W \text{ iff }  
 \begin{gathered}
   \dom(W') \supseteq \dom(W) \land \\
   \forall r \in dom(W) \ldotp W'(r) \future W(r)
 \end{gathered}
\]
Future regions:
\[
  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \text{ iff } 
    (\phi_2,H_2) = (\phi_1,H_1) \land (s_1,s_2 \in \phi_1)
\]
\begin{definition}[Heap satisfaction/erasure]
\[
\heapSat[\hs]{n}{W} \text{ iff }
\begin{gathered}
  \exists R \dom(W) \rightarrow \HeapSegments \ldotp hs = \biguplus_{r \in \dom(W)} R(r) \\
  \text{and}\\
  \forall r \in \dom(W) \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).H(W(r).s)(\roll\; W)
\end{gathered}
\]
Where $\iota.s$ and $\iota.H$ project out the state and the predicate of the region respectively.
\end{definition}

@@

Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
\begin{align*}
  \observations \quad : & \quad  \Worlds \monfun \UPred{\Regs \times \HeapSegments} \\
  \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\
                           & \quad (\forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
                           & \qquad \Rightarrow \exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W}
\end{align*}
Removed the condition $\npair[n-i]{\heap'(0)} \in \stdvr(W)$.\\

The ``good observations'' are as follows
\begin{itemize}
\item The execution does not stop in $n$ steps. 
\item The execution stops with a $\failed$.
\item The execution halts in at most $n$ steps with a heap where the island that governs address 0 is satisfied by some heap segment.
\end{itemize}
Address 0 is the designated I/O address.

@@

@@

@@

@@

\begin{lemma}[Heap satisfaction downwards closure]
\[
  \forall n,n',\heap, W \ldotp \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
\]
\end{lemma}

\begin{lemma}[Read condition downwards closure]
\[
\forall n,n',W,\start, \addrend \ldotp \readCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \readCond{n',W,\start,\addrend}
\]
\end{lemma}
\begin{lemma}[Write condition downwards closure]
\[
\forall n,n',W,\start, \addrend \ldotp \writeCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \writeCond{n',W,\start,\addrend}
\]
\end{lemma}

>>>>

\newcommand{\FS}{\var{FS}}

@@

The ticket dispenser test program uses $\malloc$. To be able to reason about programs that use $\malloc$, we assume that it follows the following specification. If you provide it with one argument, namely the size of the allocation you wish and a return capability, then at some point $\pcreg$ will contain an executable capability based on this return capability and the return register will contain a $\rwx$ capability for the freshly allocated memory. 
\begin{specification}[Malloc v.2]
  \begin{align*}
    &\forall \Phi \in \Mems \ldotp \forall \hsfoot, \hsframe \in \HeapSegments \ldotp \\
    &\quad \forall n, \size \in \nats \ldotp \forall b,e,a,m_b,m_e \in \Addrs\ldotp\forall p \in \Perms \ldotp \forall W \in \Worlds \ldotp\\
    &\qquad \exists i, \FS \ldotp \\
    &\qquad \quad W (i) = \iota_{\malloc}^{\FS} \land \\
    &\qquad \quad \dom(\hsfoot) = [m_b,m_e] \land \\
    &\qquad \quad \memheap = \hsfoot \uplus \hsframe \land\\
    &\qquad \quad \heapSat[\hsfoot]{n}{W} \land \\
    &\qquad \quad \memreg(r_1) = \size \land \\
    &\qquad \quad \size \geq 0 \land \\
    &\qquad \quad \memreg(r_0) = (p,b,e,a) \land \\
    &\qquad \quad \memreg(\pcreg) = (\entry,m_b,m_e,m_b) \\
    &\qquad \quad \Rightarrow \\
    &\qquad \qquad\exists \Phi' \in \Mems \ldotp \exists \hsfoot' \in \HeapSegments\ldotp\\
    &\qquad \qquad \quad \exists j,k \in \nats \ldotp \exists b',e'\in \Addrs \ldotp \exists W' \in \Worlds \ldotp \\
    &\qquad \qquad \qquad \Phi \step[j]\Phi' \land \\
    &\qquad \qquad \qquad \memheap[\Phi']=\hsfoot' \uplus \hsframe \land\\
    &\qquad \qquad \qquad W' \future W \land \\
    &\qquad \qquad \qquad W(i) = \iota_{\malloc}^{\FS \setminus [b',e']} \land \\
    &\qquad \qquad \qquad [b',e'] \subseteq \FS  \land \\
    &\qquad \qquad \qquad \heapSat[\hsfoot']{n-j}{W'} \land \\
    &\qquad \qquad \qquad \memreg[\Phi'](\pcreg) = (\updatePcPerm{p},b,e,a) \land \\
    &\qquad \qquad \qquad \size = e'-b' \\
    &\qquad \qquad \qquad \memreg[\Phi'](r_1) = (\rwx,b',e',b') \land \\
    &\qquad \qquad \qquad \forall r \in \RegName \setminus \{\pcreg,r_1\} \ldotp \memreg[\Phi'](r) = 0
\end{align*}
\end{specification}
The specification uses the mallog island defined as
\[
\begin{gathered}
  \iota_{\malloc}^{\FS} \defeq (\FS, \phi_{\malloc}, H_{\malloc}) \\
  \FS \subseteq \Addrs \\
  \forall \FS_1, \FS_2 \ldotp (\FS_1,\FS_2) \in \phi_{\malloc} \text{ iff } \FS_1 \supseteq \FS_2 \\
  H \; W \; \FS = {\npair{\hs} \mid \dom(\hs) = \FS}
\end{gathered}
\]
\lau{I am not satisfied with the pc in the premise. The exact capability necessary very much depent on the implementation (this is mostly a note to myself.)}
\dominique{I agree with the remark about the pc.  Perhaps leave the form of the
  pc unspecified as it is not important for malloc users?}
\dominique{Resulting frame should be equal to original.}
\dominique{Allocated pointer should have the right size.}
\dominique{Other registers unmodified could be an alternative. (Caller-save/Callee-save etc.)}

@@

\section{Logical Relation}
\label{sec:logical-relation}
\subsection{Worlds}
\[
  \Worlds = \{ W \in RegionNames \parfun \Regions \mid \dom(W) \text{ finite } \}
\]
$\States$ is a fixed set with the necessary states.

\begin{align*}
  \Regions = \{ r = (s, \phi, H) \mid &  s \in \States \land \phi \subseteq States^2 \land \\ 
                                      &  H : \States \rightarrow \StorePred \land \\ 
                                      &  \phi \text{ reflexive and transitive} \}
\end{align*}
For convenience, we write $r.s$, $s.\phi$, and $s.H$ for a the projection of a regions state, transition system, and store predicate, respectively
\[
  \StorePred = \widehat{\Worlds} \monnefun \UPred{\HeapSegments}
\]

\[
\roll : \frac{1}{2} \cdot \Worlds \cong \widehat{\Worlds}
\]
Future Worlds:
\[
 W' \future W \text{ iff }  
 \begin{gathered}
   \dom(W') \supseteq \dom(W) \land \\
   \forall r \in dom(W) \ldotp W'(r) \future W(r)
 \end{gathered}
\]
Future regions:
\[
  (s_2,\phi_2,H_2) \future (s_1,\phi_1,H_1) \text{ iff } 
    (\phi_2,H_2) = (\phi_1,H_1) \land (s_1,s_2) \in \phi_1
\]
\begin{definition}[Heap satisfaction/erasure]
\[
\heapSat[\hs]{n}{W} \text{ iff }
\exists R \ldotp 
\begin{gathered}
  \dom(W) \rightarrow \HeapSegments \ldotp hs = \biguplus_{r \in \dom(W)} R(r) \\
  \text{and}\\
  \forall r \in \dom(W) \ldotp \forall n' < n \ldotp \npair[n']{R(r)} \in W(r).H(W(r).s)(\roll\; W)
\end{gathered}
\]
\end{definition}

\begin{lemma}[Heap satisfaction uniformity]
\[
  \forall n,n',\heap, W \ldotp \heapSat{n}{W} \land n' < n \Rightarrow \heapSat{n'}{W}
\]
\end{lemma}

@@

\begin{lemma}
  \[
    \forall n' < n \ldotp \forall W \ldotp \forall w \ldotp \npair{w} \in \stdvr(W) \Rightarrow \npair[n']{w} \in \stdvr(W)
  \]
\end{lemma}

\begin{lemma}
  \[
    \forall n \ldotp \forall W' \future W \ldotp \forall w \ldotp \npair{w} \in \stdvr(W) \Rightarrow \npair{w} \in \stdvr(W')
  \]
\end{lemma}

Consider what the ``good observations'' should be? Before we do that - what should the end result of a computation be?
\begin{align*}
  \observations \quad : & \quad  \Worlds \monfun \UPred{\Regs \times \HeapSegments} \\
  \observations (W) \defeq & \{ \npair{(\reg,\hs)} \mid \\
                           & \quad (\forall \heap_f, \heap', i \leq n \ldotp (\reg,\hs \uplus \heap_f) \step[i] (\halted,\heap')  \\
                           & \qquad \Rightarrow \exists W' \future W \ldotp\exists \hs' \ldotp \heap' = \hs' \uplus \heap_f \land \heapSat[\hs']{n-i}{W'}
\end{align*}
Removed the condition $\npair[n-i]{\heap'(0)} \in \stdvr(W)$.\\

The ``good observations'' are as follows
\begin{itemize}
\item The execution does not stop in $n$ steps. 
\item The execution stops with a $\failed$.
\item The execution halts in at most $n$ steps with a heap where the island that governs address 0 is satisfied by some heap segment.
\end{itemize}
Address 0 is the designated I/O address.
\\ \lau{The two first conditions where implied by the last one, so they were removed. It turned out to cause troubles to special case on address 0, so now the entire heap has to satisfy a future world (which is a bit worrying as only the i/o address is observable).}

\begin{lemma}
  \[
    \forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \forall \hs \ldotp \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair[n']{(\reg,\hs)} \in \observations(W)
  \]
\end{lemma}

\begin{lemma}
  \[
    \forall n \ldotp \forall W \future W' \ldotp \forall \reg \ldotp \forall \hs \ldotp \npair{(\reg,\hs)} \in \observations(W) \Rightarrow \npair{(\reg,\hs)}(W')
  \]
\end{lemma}

@@

\begin{lemma}
  \[
    \forall n' < n \ldotp \forall W \ldotp \forall \reg \ldotp \npair{\reg} \in \stdrr(W) \Rightarrow \npair[n']{\reg} \in \stdrr(W)
  \]
\end{lemma}

\begin{lemma}
  \[
    \forall n \ldotp \forall W' \future W \ldotp \forall \reg \ldotp \npair{\reg} \in \stdrr(W) \Rightarrow \npair{\reg} \in \stdrr(W')
  \]
\end{lemma}

@@

\begin{lemma}
  \[
    \forall n' < n \ldotp \forall W \ldotp \forall c \ldotp \npair{c} \in \stdkr(W) \Rightarrow \npair[n']{c} \in \stdkr(W)
  \]
\end{lemma}

\begin{lemma}
  \[
    \forall n \ldotp \forall W' \future W \ldotp \forall c \ldotp \npair{c} \in \stdkr(W) \Rightarrow \npair{c} \in \stdkr(W')
  \]
\end{lemma}

@@

\begin{lemma}
  \[
    \forall n' < n \ldotp \forall W \ldotp \forall \pc \ldotp \npair{\pc} \in \stder(W) \Rightarrow \npair[n']{\pc} \in \stder(W)
  \]
\end{lemma}

\begin{lemma}
  \[
    \forall n \ldotp \forall W' \future W \ldotp \forall \pc \ldotp \npair{c} \in \stder(W) \Rightarrow \npair{c} \in \stder(W')
  \]
\end{lemma}

@@

\begin{lemma}[Read condition uniformity]
\[
\forall n,n',W,\start, \addrend \ldotp \readCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \readCond{n',W,\start,\addrend}
\]
\end{lemma}

\begin{lemma}[Write condition uniformity]
\[
\forall n,n',W,\start, \addrend \ldotp \writeCond{n,W,\start,\addrend} \land n' \leq n \Rightarrow \writeCond{n',W,\start,\addrend}
\]
\end{lemma}
